---
title: "Stock Analysis"
author: "Frank Wang"
date: "November 22, 2018"
output: html_document
---

## Stock Analysis Assignment {-}

Download stock data from http://finance.yahoo.com.  In this example, we retrieved
a year (from April 10, 2017 to April 6, 2018) of data of Apple Inc. (AAPL), saved as `AAPL.csv`.

We used the `read.csv` command to import data and assign them to the object
`dat`.  Be sure that the file is in the current directory; if not, use
the `setwd` command (Under Session/Set Work Directory).  The `head` command is for obtaining the first few rows of the data.  We show the stock price as a function of time using the `plot` command.  Note that the `type` is `l` (el) meaning line not `1` (one). 
```{r}
dat <- read.csv("AAPL.csv")
head(dat)
plot(as.Date(dat$Date), dat$Adj.Close, type = "l")
```

We assign the adjusted close to the vector called `close` for later 
convenience.  The number of trading days is the length of the vector, denoted
by `n`.
```{r}
close <- dat$Adj.Close
(n <- length(close))
```
In financial news, we often hear relative, or percentage changes.  A
stock's rate of return in any time period $t$ is defined as
$$
R_{t} = \frac{S_{t} - S_{t-1}}{S_{t-1}}
$$
where $S_{t}$ and $S_{t-1}$ are the closing prices
at time $t$ and $t-1$, respectively.  Below is the calculation.  
A histogram is made for the relative changes.  
```{r}
daily.diff <- close[2:n] - close[1:n-1]
rel.change <- daily.diff/close[1:n-1]
hist(rel.change, breaks = 20)
```

We can find the five-number summary and produce a box plot as below.  
```{r}
summary(rel.change)
boxplot(rel.change, horizontal = TRUE)
```

To find the mean and standard deviation, use the following commands.  
```{r}
(x.bar <- mean(rel.change))
(stdev <- sd(rel.change))
```
The number of trading day is `n` above
```{r}
n
```
and the number of days when the change is less than the mean is
```{r}
sum(rel.change < x.bar)
```
Express the changes as $z$-scores, and overlay the histogram of $z$
with a standard normal distribution probability density function.
```{r}
z.score <- (rel.change - mean(rel.change))/sd(rel.change)
hist(z.score, breaks = 20, freq = FALSE)
curve(dnorm(x), col = "blue", add = TRUE)
```

One financial model assumes that stocks advances with geometric Brownian motion,
$$
S_{i+1} = S_{i} + \mu S_{i} \Delta t + \sigma S_{i} \epsilon \sqrt{\Delta t}
$$
where $\epsilon$ is a random number drawn from a standardized normal distribution.  
Use \$168.38 as the initial price (the closing value on April 6, 2018),
we simulate a possible outcome for the next year.  You should re-run this 
several times to observe the possible outcomes.  
```{r}
S <- c()
close[n]
S[1] <- close[n]
for (i in 1:n){
  S[i+1] <- S[i] + x.bar*S[i]*1 + stdev*S[i]*rnorm(1)*sqrt(1)
}
plot(S, type = "l")
```


